---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# here

<!-- badges: start -->
[![Travis-CI Build Status](https://travis-ci.org/r-lib/here.svg?branch=master)](https://travis-ci.org/r-lib/here)
<!-- badges: end -->

The goal of the here package is to enable easy file referencing. 

## Installation

Install the released version of here from CRAN:
```{r, eval = FALSE}
install.packages("here")
```

Or install the development version from GitHub with:
```{r, eval = FALSE}
devtools::install_github("r-lib/here")
```

## Usage

The here package creates path relative to the top-level directory. The package displays the top-level of the current project on load or any time you call here():

```{r, message = TRUE}
library(here)
here()
```

You can build a path relative to the top-level directory in order to read or write a file:

```{r, eval = FALSE}
write.csv(iris, here("data", "iris.csv"))
```

This works, regardless of where the associated source file lives inside your project. This is especially helpful if you use RMarkdown with the default behavior of "working directory = directory where this file lives."

These paths will also "just work" during interactive development, without incessant fiddling with the working directory of your IDE’s R process.
