---
title: "here"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{here}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = !identical(Sys.getenv("IN_PKGDOWN"), "true")
)

project_path <- system.file("demo-project", package = "here")

knitr::opts_knit$set(root.dir = project_path)
```

The here package enables easy file referencing by using the top-level directory of a file project to easily build file paths.
This is in contrast to using `setwd()`, which is fragile and dependent on the way you order your files on your computer.
Read more about project-oriented workflows:

- What They Forgot to Teach You About R: ["Project-oriented workflow"](https://rstats.wtf/project-oriented-workflow.html) chapter by Jenny Bryan and Jim Hester

- ["Project-oriented workflow"](https://www.tidyverse.org/blog/2017/12/workflow-vs-script/) blog post by Jenny Bryan

- R for data science: ["Workflow: projects"](https://r4ds.had.co.nz/workflow-projects.html) chapter by Hadley Wickham

## Basic functionality

For demonstration, this article uses a data analysis project that lives in `` `r project_path` `` on my machine.
This is the *project root*.
The path will most likely be different on your machine, the here package helps deal with this situation.

The project has the following structure:

```{r echo = FALSE}
fs::dir_tree()
```

To start working on this project in RStudio, open the `demo-project.Rproj` file.
This ensures that the [working directory](https://en.wikipedia.org/wiki/Working_directory) is set to `` `r project_path` ``, the project root.
Opening only the `.R` or the `.Rmd` file may be insufficient!

Other development environments may have a different notion of a project.
Either way, it is important that the working directory is set to the project root or a subdirectory of that path.
You can check with:

```{r eval = FALSE}
setwd(project_path)
```

```{r}
getwd()
```

(See `vignette("rmarkdown")` for an example where the working directory is set to a subdirectory on start.)

The package displays the top-level directory of the current project when the package is loaded:[^attached]

[^attached]: Technically, `library()` loads and attaches a package. The message is printed only when the package is attached; if your code only uses `here::here()`, no message is printed.

```{r}
library(here)
```

In our case, the `demo-project.Rproj` file decides what the project root is.

The top-level directory is also returned from the `here()` function:

```{r}
here()
```

One important distinction from the working directory is that this remains stable even if the working directory is changed:

```{r}
setwd("analysis")
getwd()
here()
setwd("..")
```

(I suggest to steer clear from ever changing the working directory.
This may not always be feasible, in particular if the working directory is changed by code that you don't control.)


## Relative paths

You can build a path relative to the top-level directory in order to build the full path to a file:

```{r}
here("data", "penguins.csv")
readr::read_csv(here("data", "penguins.csv"))
```

This works regardless of where the associated source file lives inside your project.
With `here()`, the path will always be relative to the top-level project directory.

`here()` works very similarly to `file.path()` or `fs::path()`, you can pass path components or entire subpaths:

```{r}
here("data/penguins.csv")
```

As seen above, `here()` returns absolute paths (starting with `/`, `<drive letter>:\` or `\\`).
This makes it safe to pass these paths to other functions, even if the working directory is changed along the way.

As of version 1.0.0, absolute paths passed to `here()` are returned unchanged.
This means that you can safely use both absolute and project-relative paths in `here()`.

```{r}
data_path <- here("data")
here(data_path)
here(data_path, "penguins.csv")
```


## Situation report

The `dr_here()` function explains the reasoning behind choosing the project root:

```{r}
dr_here()
```

The `show_reason` argument can be set to `FALSE` to reduce the output to one line:

```{r}
dr_here(show_reason = FALSE)
```


## Change working directory

Unload and reload package

## Fix project root

`set_here()`

## Under the hood: rprojroot
